Very simplified likelihood

### 2024-02-22 inst/very_simplified_likelihood.Rmd
tidyup, remove old function defs

```{r,label=definefsingle}
f_single_old <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
       out <- log(B-1) - sum(log(B + (n-r-1):(n-1)))
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

f_single_old2 <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
      out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
      out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
   } else {  # not-log 
      out <- (B-1)/prod(B+(n-r-1):(n-1))
   }	
   return(out)
}

@@

MLE <- function(r,n){
    d <- 1e-6
    optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
}
MLE(0,9)
MLE(3,9)
MLE(3,7)
```

@@

```{r}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category=c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

```{r}
a <- seq(from=0.2,by=0.01,to=0.8)
wp <- category=='pure'
wa <- category=='applied'
```

@@

```{r}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

```{r}
supp2_ed <- function(vec,place1,place2,runners1,runners2){
  out <- 0
  M1 <- cbind(place1,runners1)
  M2 <- cbind(place2,runners2)
  for(i in seq_len(nrow(M1))){
    out <- out + f_vec_arn(vec[1],r=M1[i,1]-1,n=M1[i,2]-1,log=TRUE)
  }
  for(i in seq_len(nrow(M2))){
    out <- out + f_vec_arn(vec[2],r=M2[i,1]-1,n=M2[i,2]-1,log=TRUE)
  }
  return(out)
}
```

>>>>

```{r,label=definefsingle}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
      out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
   } else {  # not-log 
      out <- (B-1)/prod(B+(n-r-1):(n-1))
   }	
   return(out)
}

@@

MLE <- function(r,n){
    d <- 1e-6
    optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
}
MLE(0,9)
MLE(3,9)
MLE(3,7)

MLE_analytic <- function(r,n){
uniroot(
  f=function(B){psigamma(B-n+r-1)-psigamma(B+n)+1/(1-B)},
  interval=c(1.0001,1000)
  )
}

```

@@

```{r focaleducation}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category=c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

```{r seppurefromapplied}
a <- seq(from=0.2,by=0.01,to=0.8)
wp <- category=='pure'
wa <- category=='applied'
```

@@

```{r dopurefurst}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

```{r definesupp2ed}
supp2_ed <- function(vec,place1,place2,runners1,runners2){
  out <- 0
  M1 <- cbind(place1,runners1)
  M2 <- cbind(place2,runners2)
  for(i in seq_len(nrow(M1))){
    out <- out + f_vec_arn(vec[1],r=M1[i,1]-1,n=M1[i,2]-1,log=TRUE)
  }
  for(i in seq_len(nrow(M2))){
    out <- out + f_vec_arn(vec[2],r=M2[i,1]-1,n=M2[i,2]-1,log=TRUE)
  }
  return(out)
}
```
